---
title: "Reproduced analysis of FARS data"
author: "Janalee Thompson"
date: "November 8, 2016"
output: word_document
---

## Results
Percentages of drivers testing positive by drug type, sex, and year group

```{r, echo = FALSE, warning = FALSE, error = FALSE}
library(tidyverse)
clean_fars %>%
  mutate(year_cat = cut(year, breaks = c(1999, 2002, 2006, 2010),
                        labels = c("1999-2002", "2003-2006",
                                   "2007-2010"),
                        include.lowest = TRUE, right = TRUE)) %>%
  
  filter(!is.na(sex)) %>%
  group_by(drug_type, sex, year_cat) %>%
  
  summarize(n_non_missing = sum(!is.na(positive_for_drug)),
            positive_test = sum(positive_for_drug, na.rm = TRUE),
            perc_positive = round(100 * positive_test / n_non_missing, 1)) %>%
  
  select(drug_type, sex, year_cat, perc_positive) %>%
  
  unite(sex_year_cat, sex, year_cat) %>%
  spread(sex_year_cat, perc_positive) %>%
  knitr::kable(col.names = c("Drug type", "F 1999-2002",
                             "F 2003-2006", "F 2007-2010",
                             "M 1999-2002", "M 2003-2006",
                             "M 2007-2010"))
```

Figure 1: Prevalence of nonalcohol drugs in fatally injured drivers by year and age group

```{r, echo = FALSE, warning = FALSE, error = FALSE}

library(ggplot2)
library(ggthemes)
library(gridExtra)
library(tidyverse)
library(dplyr)

clean_fars2 <- clean_fars %>%
  dplyr::filter(!is.na(positive_for_drug),
         (!is.na(agecat))) %>%
  dplyr::filter(drug_type != "Alcohol") %>%
  dplyr::group_by(unique_id) %>%
  summarize(positive_for_drug = any(positive_for_drug),
            agecat = first(agecat),
            year = first(year)) %>%
  ungroup() %>%
  dplyr::group_by(agecat, year) %>%
  dplyr::summarize(perc_pos_drug = mean(positive_for_drug) * 100) 

plot1 <- ggplot(clean_fars2, aes(x = year, y = perc_pos_drug, shape = agecat)) +
            geom_point(alpha = 0.5) +
            geom_line() +
            scale_shape_discrete(name = "Age") +
            scale_x_continuous(name = "Year") + 
            scale_y_continuous(name = "Positive for Nonalcoholic Drugs %", breaks = c(0, 10, 20, 30), labels = c("0", "10", "20", "30")) +
  theme_few()
  plot1
```


Figure 2: Prevalence of nonalcohol drugs in fatally injured drivers by year and drug type
```{r, echo = FALSE, warning = FALSE, error = FALSE}
  
plot2 <- dplyr::group_by(clean_fars, drug_type, year) %>%
    dplyr::filter(!is.na(positive_for_drug),
         (!is.na(drug_type))) %>%
    dplyr::filter(drug_type != "Alcohol") %>%
  dplyr::summarize(perc_pos_drug = mean(positive_for_drug) * 100) %>%
ggplot(aes(x = year, y = perc_pos_drug, 
                                 shape = drug_type)) +
            geom_point(alpha = 0.5) +
            geom_line() +
            scale_shape_discrete(name = "Drug Type") +
            scale_x_continuous(name = "Year") + 
            scale_y_continuous(name = "Positive for Drugs %", breaks = c(0, 5, 10), labels = c("0", "5", "10")) +
  theme_few()
  plot2
```

Figure 3: Prevalence of cannabinoid drugs in fatally injured drivers by year and age group
```{r, echo = FALSE, warning = FALSE, error = FALSE}
  clean_fars4 <- clean_fars 
plot3 <- dplyr::group_by(clean_fars4, agecat, year) %>%
    dplyr::filter(!is.na(positive_for_drug),
         (!is.na(agecat))) %>%
    dplyr::filter(drug_type == "Cannabinoid") %>%
  dplyr::summarize(perc_pos_drug = mean(positive_for_drug) * 100) 
ggplot(plot3, aes(x = year, y = perc_pos_drug, shape = agecat)) +
            geom_point(alpha = 0.5) +
            geom_line() +
            scale_shape_discrete(name = "Age") +
            scale_x_continuous(name = "Year") + 
            scale_y_continuous(name = "Positive for Cannabinoid %", breaks = c(0, 5, 10, 15, 20), labels = c("0", "5", "10", "15", "20")) +
  theme_few()
```

Prevalence of drugs in fatally injured drivers for 1999 and 2010 by drug type
```{r, echo = FALSE, warning = FALSE, error = FALSE}

per_cis <- function(x, n){
 prop_est <- x/n
  se <- sqrt(prop_est*(1-prop_est)/n)
  lb <- prop_est - 1.96 * se
  ub <- prop_est + 1.96 * se
  prop_new <- paste0(round((prop_est * 100), 1), "% ", 
                     "(",round((lb * 100), 1),"%", ", ", round((ub * 100), 1),"%)")
  return(prop_new)
}

clean_fars %>%
filter(year == 2010 | year == 1999) %>%
  group_by(drug_type, year) %>%
dplyr::summarize(n = sum(!is.na(positive_for_drug)),
        x = sum(positive_for_drug, na.rm = TRUE)) %>%
  mutate(value = per_cis(x,n)) %>%
  select(drug_type, year, value) %>%
  spread(key = year, value = value) %>% 
  rename(`Drug type` = drug_type) %>%
  knitr::kable()
```

Statistics for testing for trend in prevalence of drugs over study years by drug type using Cochran-Armitage trend test
```{r, echo = FALSE, warning = FALSE, error = FALSE}
  
drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
                  "Stimulant", "Cannabinoid", "Other")
   drug_trend_tests_ca <- lapply(drug_list, test_trend_ca)
   drug_trend_tests_ca <- dplyr::bind_rows(drug_trend_tests_ca) %>%
     dplyr::mutate(drug = drug_list) %>%
     dplyr::select(drug, Z, p.value)
   drug_trend_tests_ca %>% knitr::kable()
```

Statistics for testing for trend in prevalence of drugs over study years by drug type using Wald test of logistic regression coefficient for “year”
```{r, echo = FALSE, warning = FALSE, error = FALSE}
      drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
                  "Stimulant", "Cannabinoid", "Other")
   drug_trend_tests_log_reg <- lapply(drug_list, test_trend_log_reg)
   drug_trend_tests_log_reg <- dplyr::bind_rows(drug_trend_tests_log_reg) %>%
     dplyr::mutate(drug = drug_list) %>%
     dplyr::select(drug, Z, p.value)
   drug_trend_tests_log_reg %>% 
     knitr::kable()
```

